Step 13: Marginal Effects Analysis with Multiple Imputation

### 2025-09-08 reanalysis-13-marginal_effects.qmd
ame bug

  # Display results if available
  if (!is.null(test_results[[sg_name]])) {
    mfx_summary <- test_results[[sg_name]]
    if ("ever_lapse_binary" %in% rownames(mfx_summary)) {
      ame <- mfx_summary["ever_lapse_binary", "AME"]
      se <- mfx_summary["ever_lapse_binary", "SE"]
      cat(sprintf("  AME = %.4f (SE = %.4f)\n", ame, se))
    }
  }
}

@@

      if (is.null(sg$subset_condition)) {
        # Overall marginal effect
        mfx <- margins(model, design = design, variables = "ever_lapse_binary")
        mfx_summary <- summary(mfx)
      } else {
        # Subset marginal effect
        subset_data <- subset(imp_data, eval(sg$subset_condition))

@@

      # Store results
      if (!is.null(mfx_summary) && "ever_lapse_binary" %in% rownames(mfx_summary)) {
        all_mfx_results[[sg_name]][[i]] <- list(
          ame = mfx_summary["ever_lapse_binary", "AME"],
          se = mfx_summary["ever_lapse_binary", "SE"],
          n = ifelse(is.null(sg$subset_condition), 
                    nrow(imp_data),
                    nrow(subset(imp_data, eval(sg$subset_condition)))),
          success = TRUE
        )
      } else {
        all_mfx_results[[sg_name]][[i]] <- list(success = FALSE)
      }
    }

@@

```{r summary-table}
# Create summary table
summary_df <- do.call(rbind, lapply(names(pooled_mfx), function(sg_name) {
  res <- pooled_mfx[[sg_name]]
  if (res$success) {
    data.frame(
      Subgroup = res$name,
      AME = sprintf("%.4f", res$ame),
      SE = sprintf("%.4f", res$se),
      `95% CI` = sprintf("(%.4f, %.4f)", res$ci_lower, res$ci_upper),
      `P-value` = ifelse(res$p_value < 0.001, "<0.001", 
                         sprintf("%.3f", res$p_value)),
      `Avg N` = sprintf("%.0f", res$avg_n),
      `N Imp` = res$n_imputations,
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  }
}))

# Display table - simplified to avoid kable errors
if (nrow(summary_df) > 0) {
  # Use basic kable without complex styling that can fail with long data
  print(kable(summary_df,
        caption = "Pooled Average Marginal Effects of Lapsing by Subgroup",
        align = c("l", "r", "r", "c", "r", "r", "r"),
        digits = 4))
} else {
  cat("No results to display in summary table\n")
}

# Save to CSV
write.csv(summary_df, 
          file.path(marginal_effects_dir, "marginal_effects_summary.csv"),
          row.names = FALSE)
log_message("Saved summary table to CSV")
```

>>>>

  # Display results if available
  if (!is.null(test_results[[sg_name]])) {
    mfx_summary <- test_results[[sg_name]]

    # Debug output for first subgroup
    if (sg_name == names(subgroups)[1]) {
      cat("  Debug - margins output structure:\n")
      cat("    Rows:", paste(rownames(mfx_summary), collapse=", "), "\n")
      cat("    Cols:", paste(colnames(mfx_summary), collapse=", "), "\n")
      if ("factor" %in% colnames(mfx_summary)) {
        cat("    Factor column:", paste(mfx_summary$factor, collapse=", "), "\n")
      }
    }

    # Margins output has factor name in 'factor' column, not rownames
    # Look for ever_lapse_binary in the factor column
    if ("factor" %in% colnames(mfx_summary)) {
      lapse_rows <- grep("ever_lapse_binary", mfx_summary$factor)
      if (length(lapse_rows) > 0) {
        # Use the first matching row
        lapse_row <- lapse_rows[1]
        ame <- mfx_summary[lapse_row, "AME"]
        se <- mfx_summary[lapse_row, "SE"]
        cat(sprintf("  AME = %.4f (SE = %.4f)\n", ame, se))
      } else {
        cat("  WARNING: Could not find ever_lapse_binary in factor column\n")
      }
    } else {
      cat("  WARNING: No 'factor' column in margins output\n")
    }
  }
}

@@

      tryCatch({
        if (is.null(sg$subset_condition)) {
          # Overall marginal effect
          mfx <- margins(model, design = design, variables = "ever_lapse_binary")
          mfx_summary <- summary(mfx)
        } else {
          # Subset marginal effect
          subset_data <- subset(imp_data, eval(sg$subset_condition))

@@

        # Store results - margins output has factor name in 'factor' column
        if (!is.null(mfx_summary)) {
          # Debug: show what we got for first imp and first subgroup
          if (i == 1 && sg_name == names(subgroups)[1]) {
            cat("  Debug: mfx_summary structure:\n")
            cat("    Cols:", paste(colnames(mfx_summary), collapse=", "), "\n")
            if ("factor" %in% colnames(mfx_summary)) {
              cat("    Factor values:", paste(mfx_summary$factor, collapse=", "), "\n")
            }
          }

          # Look for ever_lapse_binary in the factor column
          if ("factor" %in% colnames(mfx_summary)) {
            lapse_rows <- grep("ever_lapse_binary", mfx_summary$factor)
            if (length(lapse_rows) > 0) {
              # Use the first matching row (should only be one)
              lapse_row <- lapse_rows[1]
              all_mfx_results[[sg_name]][[i]] <- list(
                ame = mfx_summary[lapse_row, "AME"],
                se = mfx_summary[lapse_row, "SE"],
                n = ifelse(is.null(sg$subset_condition), 
                          nrow(imp_data),
                          nrow(subset(imp_data, eval(sg$subset_condition)))),
                success = TRUE
              )
            } else {
              all_mfx_results[[sg_name]][[i]] <- list(
                success = FALSE,
                error = "Could not find ever_lapse_binary in factor column"
              )
            }
          } else {
            # Fallback: if no factor column, try using first row
            if (nrow(mfx_summary) == 1) {
              all_mfx_results[[sg_name]][[i]] <- list(
                ame = mfx_summary[1, "AME"],
                se = mfx_summary[1, "SE"],
                n = ifelse(is.null(sg$subset_condition), 
                          nrow(imp_data),
                          nrow(subset(imp_data, eval(sg$subset_condition)))),
                success = TRUE
              )
            } else {
              all_mfx_results[[sg_name]][[i]] <- list(
                success = FALSE,
                error = "No 'factor' column and multiple rows in margins output"
              )
            }
          }
        } else {
          all_mfx_results[[sg_name]][[i]] <- list(
            success = FALSE,
            error = "No data in subset"
          )
        }
      }, error = function(e) {
        all_mfx_results[[sg_name]][[i]] <- list(
          success = FALSE,
          error = e$message
        )
        if (i == 1) {
          cat(paste0("    ERROR in ", sg_name, ": ", e$message, "\n"))
        }
      })
    }

@@

```{r summary-table}
# Create summary table
if (length(pooled_mfx) > 0) {
  summary_rows <- lapply(names(pooled_mfx), function(sg_name) {
    res <- pooled_mfx[[sg_name]]
    if (!is.null(res) && !is.null(res$success) && res$success) {
      data.frame(
        Subgroup = res$name,
        AME = sprintf("%.4f", res$ame),
        SE = sprintf("%.4f", res$se),
        `95% CI` = sprintf("(%.4f, %.4f)", res$ci_lower, res$ci_upper),
        `P-value` = ifelse(res$p_value < 0.001, "<0.001", 
                           sprintf("%.3f", res$p_value)),
        `Avg N` = sprintf("%.0f", res$avg_n),
        `N Imp` = res$n_imputations,
        stringsAsFactors = FALSE,
        check.names = FALSE
      )
    } else {
      NULL
    }
  })

  # Remove NULL entries and combine
  summary_rows <- summary_rows[!sapply(summary_rows, is.null)]

  if (length(summary_rows) > 0) {
    summary_df <- do.call(rbind, summary_rows)

    # Display table - simplified to avoid kable errors
    print(kable(summary_df,
          caption = "Pooled Average Marginal Effects of Lapsing by Subgroup",
          align = c("l", "r", "r", "c", "r", "r", "r"),
          digits = 4))

    # Save to CSV
    write.csv(summary_df, 
              file.path(marginal_effects_dir, "marginal_effects_summary.csv"),
              row.names = FALSE)
    log_message("Saved summary table to CSV")
  } else {
    cat("No successful results to display in summary table\n")
    log_message("No successful pooled results for summary table")
  }
} else {
  cat("No pooled results available for summary table\n")
  log_message("No pooled results found")
}
```
